GPU-friendly Stroke Expansion

### 2024-04-25 paper.tex
Checkpoint writing
Cusp, introduction, prior work, add appendix for Geometric Hermite interpolation.

TODO: brief description of stroke to fill conversion.

@@

TODO: the introduction should summarize the paper's contributions and the outline of the rest of the paper.
TODO: discussion of prior art.

\begin{itemize}
    \item Nehab 2020
    \item Kilgard 2020: Polar stroking (discussion of error metric can move here)
    \item Yzerman 2020: combined offset + flatten using quadratics
    \item Blog post of Dragoș?
    \item Curve-to-curve offsetting? Tiller and Hanson (which is good for quadratics, terrible for cubics), Shape control (pretty good but has robustness issues). Hain (misses error tolerance in cusp case, as described in Yzerman). The "Comparing Offset Curve Approximation Methods" paper is flawed.
\end{itemize}

@@

Given tangent angles relative to the chord, finding the Euler spiral segment that minimizes total curvature variation is a form of geometric Hermite interpolation. There are a number of published solutions to this problem, all involving nontrivial numerical solving techniques: gradient descent\cite{Kimia2003}, bisection\cite{Walton2009}, or Newton iteration\cite{Connor2014}. Our approach is much more direct. We observe that the Euler spiral parameters are a smooth function of the two input parameters, and determine a 2D Taylor's series for this function. Many of the terms are zero because of symmetry. Preserving terms up to 7th order provides excellent accuracy (TODO: quantify more carefully) with modest computational expense.

TODO: A good cite for this is \cite{Reif2021}. Our solution is similar in flavor, but is more precise at modest cost as it contains more polynomial terms.

TODO: the technique used is a refinement of what was sketched in the Python notebook attached to the "cleaner parallel curves with Euler spirals" blog post. It's basically a 2D Taylor's series. Previously we had the secant method from section 8.2 of my thesis; not sure whether we want to cite that in the submission for anonymity reasons. It's functionally pretty similar to the \cite{Connor2014} reference.

@@

A significant concern with rendering stroked outlines of cubic Béziers is dealing with cusps and near-cusps in the input curve. A Bézier curve is expressed in parametric form, and the derivative with respect to the parameter can be zero or nearly so, causing serious numerical problems for many stroke expansion algorithms. In the limit, as the Bézier curve describes a semi-cubical parabola, the curvature can become infinite.

@@

@@

>>>>

Stroke expansion is the process of generating an outline that, when filled, produces the \emph{stroke} of a given input path. Informally, a line of the same width as the stroke is swept along the path, normal to it, and also decorated with joins and caps. Though stroke rendering is referenced in many 2D graphics standards, it has lacked a precise formal definition until recent work, primarily \cite{Nehab2020} and \cite{Kilgard2020}.

@@

A great number of rendering techniques for strokes have been proposed. Some are \emph{local,} in that they break the stroke up into smaller closed pieces, the union of which forms an approximation to the stroke. Included in this category is polar stroking \cite{Kilgard2020}. Others are based on stroke expansion, generating an explicit outline, which is then filled to produce the stroke rendering. Such techniques are further divided into \emph{flattening} approaches, where the outline is approximated by a polyline, and \emph{curve-based} approaches, exemplified by the approach in \cite{Nehab2020}. In both case, the outline contains approximations of parallel curves of the input curve segments, and, when targeting strong correctness, evolutes as well. For a more detailed survey of existing stroke rendering techniques, the reader is directed to Section 4 of \cite{Nehab2020}.

The subproblem of approximating parallel curves has also spawned extensive literature, none of which is entirely satisfying, especially when it comes to algorithms that can be efficiently evaluated on GPU. An example of a reasonably good algorithm that computes flattened parallel curves is \cite{Yzerman2020}, as used in the Blend2D rendering library. For producing curved outlines, the Tiller and Hanson algorithm\cite{Tiller1984} is commonly implemented and cited, but its performance is quite poor when applied to cubic Béziers. The quadratic Bézier version is adequate, and forms the basis of the algorithm in \cite{Nehab2020}. A variant is also used in Skia\cite{Skia}; instead of lowering the cubic Bézier to a quadratic approximation and then computing the parallel curve, Skia approximates the parallel curve directly with a quadratic Bézier, then measures the error to determine whether further subdivision is necessary.

This paper presents a comprehensive solution to stroke expansion, well suited to GPU implementation. It can produce both flattened polylines or an approximation consisting of circular arcs. The best choice depends on the capabilities of the path rendering mechanism following stroke expansion, but as we show, outlines consisting of circular arcs have many fewer segments and are correspondingly faster to produce. The algorithm is based on Euler spiral segments as an intermediate representation, with an iterative algorithm based on a straightforward error metric for conversion from cubic Béziers. We have also devised a compact binary encoding of paths, suitable for fully parallel computation of stroke outlines, while requiring minimal CPU-side processing. Our algorithm has been implemented in GPU compute shaders, integrated in a full rendering engine for vector graphics, and shows a dramatic speedup over CPU methods.

@@

Given tangent angles relative to the chord, finding the Euler spiral segment that minimizes total curvature variation is a form of geometric Hermite interpolation. There are a number of published solutions to this problem, involving nontrivial numerical solving techniques: gradient descent\cite{Kimia2003}, bisection\cite{Walton2009}, or Newton iteration\cite{Connor2014}. A more direct approach is to approximate the function as a reasonably low-order polynomial in terms of the endpoint angles. Our approach is to use a 7th order polynomial, which is more precise than the 3rd order polynomial proposed in \cite{Reif2021}, at only modest increased cost. The exact coefficients used are presented in Appendix \ref{appendix:gh}.

@@

There are two types of cusps that must be handled in the stroke expansion problem. One is when the input cubic Bézier contains a cusp (or near-cusp, as there are numerical robustness issues), and one is when the parallel curve contains cusps. This section will describe both in order.

A Bézier curve is expressed in parametric form, and the derivative with respect to the parameter can be zero or nearly so, causing serious numerical problems for many stroke expansion algorithms. In the limit, as the Bézier curve describes a semi-cubical parabola, the curvature can become infinite.

@@

\subsubsection{Cusps in parallel curves}

Even when the input curve is smooth, its parallel curve contains a cusp when the (signed) curvature equals the offset distance (the half-width of the stroke). In published techniques, dealing with these cusps is a nontrivial effort, and involves numerical methods that are not GPU friendly. In particular, detecting locations in the cubic Bézier where the curvature crosses a given quantity is a medium-degree polynomial, and in general requires numerical techniques for root finding. In the approach of \cite{Nehab2020}, the root finder not only requires a hybrid Newton/bisection method, which requires iteration, but is also recursive in that it uses the roots of a polynomial of one degree lower as a subroutine. In general, polynomials up to cubic can be considered GPU-friendly, while finding cusps in cubic Béziers requires a higher degree.

In the simplest case, weak stroke correctness with line segments as the output primitive, no additional work is needed -- the flattening algorithm for Euler spiral parallel curves will naturally generate a point within the given error tolerance of the true cusp. However, generation of arc segments and drawing the evolutes needed for strong correctness both require subdivision at the parallel curve cusps.

Fortunately, finding the cusp in the parallel curve of an Euler spiral parallel curve is a simple linear equation, and there is at most one cusp in any such segment. Arguably, Euler spirals are a reasonably good solution to problem of determining cusps, as a piecewise linear approximation to curvature, even aside from their usefulness in flattening.

@@

\appendix

\section{Geometric Hermite interpolation for Euler spiral}
\label{appendix:gh}

This appendix gives the detailed algorithm for Geometric Hermite interpolation, determining Euler spiral segment parameters given the tangent angles at the endpoints relative to the chord.

The Euler spiral is represented by $\kappa(s) = k_0 + k_1s$, where $s$ ranges from 0 to 1, ie the arc length is unit normalized. This segment is scaled, rotated, and translated into place so that its endpoints match the desired locations. Immediately, $k_0$ can be determined as $\theta_0 + \theta_1$. Therefore, the remaining parameters to compute are $k_1$ and the ratio of chord length to arc length. The latter \emph{could} be computed from $k_0$ and $k_1$ by numerical integratoiun, but it is more efficient to obtain it at the same time as $k_1$. Setting $k = \theta_0 + \theta_1$ and $\Delta = \theta_1 - \theta_0$, we have:

\[
    \setlength\arraycolsep{2pt}
    \begin{array}{rl}
    k_1 = & 6\Delta \\
     - & \Delta^3 / 70 \\
     - & \Delta^5 / 10780 \\
     + & \Delta^7 \cdot 2.769178184818219\times 10^{-7} \\
     - & k^2\Delta / 10 \\
     + & k^2\Delta^3 / 4200 \\
     + & k^2\Delta^5 \cdot 1.6959677820260655\times 10^{-5} \\
     - & k^4\Delta / 1400 \\
     + & k^4\Delta^3 \cdot 6.84915970574303\times 10^{-5} \\
     - & k^6\Delta \cdot 7.936475029053326\times 10^{-6}
    \end{array}
\]

Similarly, the formula for the chord to arc length ratio:

\[
    \setlength\arraycolsep{2pt}
    \begin{array}{rl}
    c = & 1 \\
    - & \Delta^2 / 40 \\
    + & \Delta^4 \cdot 0.00034226190482569864 \\
    - & \Delta^6 \cdot 1.9349474568904524\times 10^{-6} \\
    - & k^2 / 24 \\
    + & k^2\Delta^2 \cdot 0.0024702380951963226 \\
    - & k^2\Delta^4 \cdot 3.7297408997537985\times 10^{-5} \\
    + & k^4 / 1920 \\
    - & k^4\Delta^2 \cdot 4.87350869747975\times 10^{-5} \\
    - & k^6 \cdot 3.1001936068463107\times 10^{-6}
    \end{array}
\]

Note that the latter quantity is equal to $\mbox{sinc}(k/2)$ when $\Delta = 0$.

The coefficients were determined by numerical differentiation, using a Newton-style solver as the source of truth. The resulting formulas are extremely accurate over a wide range of inputs.

% TODO: we can definitely quantify that more precisely.
